A Historical Analysis of Quark Matter Conferences: \\ Trends, Participation, and Topic Evolution

### 2025-03-18 QM/paper.tex
keyward analysis

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/conference_venues.pdf}
\caption{Geographical distribution of Quark Matter conference venues from 2011 to 2025. The timeline displays each conference location, color-coded by continent. This visualization illustrates the global reach of the conference series and regional patterns in host selection over time.}
\label{fig:venues}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/top_keywords.pdf}
\caption{Evolution of top research topics in Quark Matter conferences from 2011 to 2025. Each cluster of bars represents a conference year, with the seven most frequent keywords from presentation titles shown with their occurrence counts. This visualization reveals shifts in the field's focus over time, highlighting emerging and declining research areas.}
\label{fig:keywords}
\end{figure}

Figure~\ref{fig:keywords} presents the evolution of research topics across Quark Matter conferences from 2011 to 2025, as revealed by keyword frequency analysis of presentation titles. Several notable trends emerge from this visualization:

The keyword "QGP" (Quark-Gluon Plasma) consistently appears among the top terms across most conference years, reflecting the central focus of the field. However, its relative prominence varies over time, suggesting shifts in how directly presentations address the QGP versus specific properties or phenomena within it.

We observe a clear evolution in experimental systems of interest. Early conferences show significant focus on "LHC" and "RHIC" accelerator facilities, while later years show more diversified keywords related to specific measurements and phenomena, indicating a maturation of the field from facility-focused to phenomenon-focused research.

Certain physics concepts show interesting temporal patterns. Terms related to "flow" and "collectivity" are prominent in middle years, corresponding to the period of intense investigation into hydrodynamic behavior in heavy-ion collisions. Later conferences show increased occurrence of terms related to "small systems" and "fluctuations," reflecting the field's expanding interest in these areas.

Theoretical concepts also show clear evolution, with early emphasis on "model" and "simulation" gradually giving way to more specific theoretical approaches and phenomena, suggesting increasing theoretical sophistication and specialization.

This keyword analysis provides a data-driven view of how research priorities in the heavy-ion physics community have evolved over the past decade and points to potential future directions as indicated by emerging keywords in recent conferences.

>>>>

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/conference_venues.pdf}
\caption{Geographical distribution of Quark Matter conference venues from 2011 to 2025. The map shows the global spread of conference locations across three continents, with numbered markers indicating the chronological sequence. Venues are color-coded by continent (blue for Europe, red for North America, and yellow for Asia), and connected by lines showing the progression between conferences. This visualization illustrates both the international nature of the conference series and the concentration of venues in certain regions.}
\label{fig:venues}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/QM_keyword_analysis.pdf}
\caption{Comprehensive analysis of research topics in Quark Matter conferences from 2011 to 2023. The top panels show the evolution of key physics concepts (left) and experimental facilities (right) as percentages of total presentations. The bottom panels display word clouds for each conference year, highlighting the most prominent terms in presentation titles after removing common stopwords. This multi-faceted visualization reveals both quantitative trends and qualitative shifts in research focus over time.}
\label{fig:keywords}
\end{figure}

Figure~\ref{fig:keywords} presents a comprehensive analysis of research topics across Quark Matter conferences from 2011 to 2023. This enhanced visualization combines quantitative trend analysis with qualitative word clouds to provide deeper insights into the field's evolution:

The upper-left panel tracks specific physics concepts over time, revealing clear temporal patterns in research focus. "QGP \& Plasma" terms maintain a consistent presence throughout the period, confirming the central role of quark-gluon plasma studies in the field. However, we observe distinct phases in other research areas: "Flow \& Collectivity" shows peak interest in the middle years (2014-2018), while "Small Systems" and "Fluctuations" gain prominence in later conferences, reflecting the community's expanding research horizons.

The upper-right panel monitors mentions of experimental facilities, documenting the field's transition from facility-focused to phenomenon-focused research. Early conferences show strong emphasis on LHC and RHIC facilities, with a gradual decline in facility mentions as the research community moved toward more specific physical phenomena and measurements. The increasing mentions of future facilities in recent years (EIC, FAIR, NICA) signals the field's forward-looking perspective and preparation for next-generation experiments.

The word clouds for individual conference years provide qualitative context for these trends, highlighting the most distinctive terms after removing common stopwords. In our analysis, we deliberately filtered out trivial keywords that appear frequently but provide little insight into research trends. Terms such as "quark," "matter," "collision," "heavy-ion," "measurement," "study," and "analysis" were removed despite their high frequency, as they represent generic terminology common to all presentations rather than indicators of specific research directions. This filtering process allows the visualization to highlight meaningful physics concepts and emerging research areas that would otherwise be obscured by these ubiquitous terms.

These visualizations reveal subtle shifts in terminology and emphasis that complement the quantitative analysis. For instance, we can observe the emergence of terms related to machine learning and computational approaches in more recent conferences, reflecting the field's adoption of new analytical techniques.

Theoretical evolution is also evident in the changing prominence of terms across years. Early conferences feature more general theoretical frameworks, while later years show increasing specificity in theoretical approaches, suggesting growing theoretical sophistication and specialization as the field matures.

This multi-faceted keyword analysis provides a data-driven view of how research priorities in the heavy-ion physics community have evolved over the past decade and points to potential future directions as indicated by emerging keywords in recent conferences.
